---
title: Visualising the Effect of Interaction Variables in Logistic Regression
subtitle: A Learning Tool
author: Taiki Sakai
job: Top Secret Astronaut
framework: io2012
highlighter: highlight.js
hitheme: tomorrow
widgets: [mathjax]
mode: selfcontained
---

## Introduction

The purpose of this shiny app is to allow students to visualise the
effect that an interaction term can have on a logistic regression. 
Interaction terms are often added to regression, but it can be difficult
to understand exactly what they are doing. The app is intended as a 
learning tool to help students better understand how an interaction term
will effect their regression model.

---

## Basic Logistic Regression

```{r, echo=FALSE, results='hide'}
library(ggplot2)
library(gridExtra)
```

```{r, echo=FALSE, fig.width=12, fig.height=4.5}
g <- ggplot( data.frame(x=c(-20,20)), aes(x)) +
      stat_function(fun=function(x) exp(3+0.5*x)/(1+exp(3+0.5*x))) +
      coord_cartesian(xlim=c(-20,20),ylim=c(-.1,1.1)) +
      labs(title='log(odds) = 3+0.5x',y='Probability')

f <- ggplot( data.frame(x=c(-20,20)), aes(x)) +
      stat_function(fun=function(x) exp(-1+2*x)/(1+exp(-1+2*x))) +
      coord_cartesian(xlim=c(-20,20),ylim=c(-.1,1.1)) +
      labs(title='log(odds) = -1+2x',y='Probability')
grid.arrange(g,f,ncol=2)
```

Recall that in basic logistic regression, our model takes the form of
 $log(odds) = a + bx$. The graph on the left above shows the probability of a 
positive result when our model is $log(odds) = 3+0.5x$. The graph on the right
shows the result when our model is changed to $log(odds) = -1 + 2x$. The result
of this change is easy for students to understand because it is the same as the
change when graphing a regular line. Increasing the x coefficient increases the slope,
and changing the intercept term shifts the graph.

---

## Logistic Regression with an Interaction Term

```{r, echo=FALSE, fig.width=12, fig.height=4.5}
h <- ggplot( data.frame(x=c(-20,20)), aes(x)) +
      stat_function(fun=function(x) exp(2.5+1.5*x)/(1+exp(2.5+1.5*x))) +
      coord_cartesian(xlim=c(-20,20),ylim=c(-.1,1.1)) +
      labs(title='log(odds) = 3-0.5z+x(0.5+z)',y='Probability',x='z=1')

j <- ggplot( data.frame(x=c(-20,20)), aes(x)) +
      stat_function(fun=function(x) exp(3.5-0.5*x)/(1+exp(3.5-0.5*x))) +
      coord_cartesian(xlim=c(-20,20),ylim=c(-.1,1.1)) +
      labs(title='log(odds) = 3-0.5z+x(0.5+z)',y='Probability',x='z=-1')
grid.arrange(h,j,ncol=2)
```

What happens if we add an interaction term, $z$? Now our function is
$log(odds) = 3-0.5z+x(0.5+z)$. This is the same as the function used on the
left graph in the previous slide with an interaction term, $z$, added. It is not
as obvious what the coefficients of -0.5 and 1 mean, and of course the graph
also depends on the value of $z$. Here we see the same function with $z$ values of
1 and -1. 

---

## The App

The shiny app will help students better understand the coefficients of the
interaction term $z$ by allowing them to adjust the parameters and see
the result of the change in real time. By playing around with different
combinations of variable changes, students can get a better sense of
how each piece affects the graph as a whole.
